### 2017-06-06 paper/documents/0.abstract.tex
- update paper
* fixed grammar

Multi-player online battle arena games are the most successful games. Every players have own specific roles in this team based game. Dota2 is the most popular for this games. This paper analyzing player roles and their behavior using supervised machine learning and also investigates the applicability of feature normalization to enhance accuracy of prediction result.We provide an in-depth discussion and novel approaches for constructing complex attributes from low-level data extracted from replay files. Using attribute evaluation techniques, we are able to reduce a larger set of candidate attributes down to a manageable number. Random Forest is the most stable and best performing classifier in our results.

>>>>

State-of-the-art, Computer games are big business, which is also reflected in the growing interest in competitive gaming. \textit{Multi-player online battle arena} games are the most successful competitive gaming in computer games industry. Every player has own specific roles in this team based game. \textit{Dota 2} is the most popular for this games. This paper analyzing player roles and their behavior using \textit{supervised machine learning} and also investigates the applicability of feature normalization to enhance the accuracy of prediction result.We provide an in-depth discussion and novel approaches for constructing complex attributes from low-level data extracted from replay files. Using attribute evaluation techniques, we are able to reduce a larger set of candidate attributes down to a manageable number. \textit{Random Forest} the decision trees algorithm is the most stable and best-performing classifier in our results. Again, \textit{Random Forest} although not the best performing classifier it is proved to be very stable and well suited to this domain.

### 2017-06-06 paper/documents/1.introduction.tex
- update paper
* fixed grammar

Computer games is a big business. Competitive gaming growing the interest in this business. The competitive gaming events, the so-called electronic sports (sPorts). State-of-the-art multi-player online batle arena (MOBA) games are the most popular and succesful games in this regard. Game tournaments always give significant awards with big prize of money. Mostly, professional player can make a living from the prize.

In this paper we analyzing the player roles using supervised machine learning (ML) , investigate the applicability of feature normalization to enhance accuracy of prediction results. We also comparing our result with previous paper. Christoph Eggert on his paper said that their most stable and performing classifier is logistic regression. We will use the same classification method with their work. The different is we will using ROpenDota to built the data set.

An approach to ML in computer games in general was proposed by Drachen et al.\cite{drachen2014skill} . They suggest using unsupervised learning algorithms, specifically k-means and \textit{Simplex Volume Maximization}, to cluster player behavior data.

Our investigation about applicability and performance of \textit{supervised machine learning} (ML) to classify player roles based on behavior in \textit{Dota 2}, the most popular MOBA game. Such information could be useful for game designer to better understand how their game design influences emergent gameplay and player behavior but also for players, both casual and professional, who want to analyze their own performance or who want to learn from the others. It could also support casters and moderators in commentating and presenting matches. Furthermore, this research might hold implications for social and other research concerned with (human behavior in) games. While ML has been applied to games and traditional sports, most works are either interested in questions like spatial behavior, trying to predict the match outcome, or otherwise trying to correlate performance to certain events or behaviors. In contrast, we aim at building a classifier that is largely independent of individual players performance and that is also not tied to the overall match outcome but that is able to identify a player's role in term of the non-formally defined roles established as common grounds within the \textit{Dota 2} or MOBA community.

This paper contributes to the state of the art in several ways: We provide an in-depth discussion and novel approaches regarding the construction of complex attributes from low level data extracted from Dota 2 replay files, together with an evaluation of these attributes with respect to different classifiers. Based on the resulting reduced set of attributes, we compare and discuss the performance of a range of supervised classification algorithms, including logistic regression, random forest decision trees, support vector machines (in combination with Sequential Minimal Optimization), naive Bayes and Bayesian networks, classifying both with a newly established larger set of player roles, as well as with a reduced set inspired by related work ~\cite{eggert2015classification}.

>>>>

Computer games is a big business. Competitive gaming growing the interest in this business. The competitive gaming events, the so-called electronic sports (sPorts). State-of-the-art, multi-player online battle arena (MOBA) games are the most popular and successful games in this regard. Game tournaments always give significant awards with the big prize of money. Mostly, the professional player can make a living from the prize.

In this paper we analyze the player roles using supervised machine learning (ML) , investigate the applicability of feature normalization to enhance the accuracy of prediction results. We also compare our result with previous paper. Christoph Eggert on his paper said that their most stable and performing classifier is the logistic regression. We will use the same classification method with their work. The different is we will using ROpenDota to build the data set.

An approach to ML in computer games, in general, was proposed by Drachen et al.\cite{drachen2014skill} . They suggest using unsupervised learning algorithms, specifically k-means and \textit{Simplex Volume Maximization}, to cluster player behavior data.

Our investigation about applicability and performance of \textit{supervised machine learning} (ML) to classify player roles based on behavior in \textit{Dota 2}, the most popular MOBA game. Such information could be useful for the game designer to better understand how their game design influences emergent gameplay and player behavior but also for players, both casual and professional, who want to analyze their own performance or who want to learn from the others. It could also support casters and moderators in commentating and presenting matches. Furthermore, this research might hold implications for social and other research concerned with (human behavior in) games. While ML has been applied to games and traditional sports, most works are either interested in questions like spatial behavior, trying to predict the match outcome, or otherwise trying to correlate performance to certain events or behaviors. In contrast, we aim at building a classifier that is largely independent of an individual players performance and that is also not tied to the overall match outcome but that is able to identify a player's role in term of the non-formally defined roles established as common grounds within the \textit{Dota 2} or MOBA community.

This paper contributes to the state of the art in several ways: We provide an in-depth discussion and novel approaches regarding the construction of complex attributes from low-level data extracted from Dota 2 replay files, together with an evaluation of these attributes with respect to different classifiers. Based on the resulting reduced set of attributes, we compare and discuss the performance of a range of supervised classification algorithms, including logistic regression, random forest decision trees, support vector machines (in combination with Sequential Minimal Optimization), naive Bayes and Bayesian networks, classifying both with a newly established larger set of player roles, as well as with a reduced set inspired by related work ~\cite{eggert2015classification}.

### 2017-06-06 paper/documents/2.background.tex
* fixed grammar

\textit{Dota 2} is team based game in \textit{Multi-player Online Battle Arena} (MOBA). It is divided by two areas, the \textit{Dire} and the \textit{Radiant}. Each team have five players. Every player will choose their hero in pre-match games. Currently, there is 113 heroes, 166 current items. \textit{Dota 2} game design is separated the heroes with three types. The first is agility heroes type, the second is strength heroes type and the last is intelligent heroes type. Every heroes have own specialty and weakness. The items can be combine into upgrade-able item which is will give more status in heroes abilities.

Hero selection is important aspect of the game. Team needs to balance the heroes with different abilities that able to fulfill certain of roles and strategies. In the beginning of the game, each player will get 650 gold to buy starting items.

\textit{Dota 2} map layout split into three lanes. Each lanes has three defensive towers. the Dire lanes and the Radiant lanes are quite different. the top of Radiant it is called off-lane , meanwhile in Dire, the top lane is called safe-lane. The bottom lane of Radiant is called safe-lane, while in Dire it is called off-lane. And the middle lane will remain same.

@@

Player roles or play styles is the way how player will choose and try to act as their roles or styles in the games. It is important to note that these roles describe a different facet of play than classic player type classifications (e.g. after Bartle \cite{bartle1996hearts}) which aim to classify expressed character traits of the players and were designed to match role-play style games. Comparison of videos, Online guides, and commentary of professional players and commentators are our sources for selection and characterization of the player roles. Player roles definitions and naming conventions will in any case differ slightly among the player base and shift over time as the game evolves, constituting another challenge for \textit{Machine Learning} applications in this area. We isolated five player roles for the main \textit{Machine Learning} task, which strike a balance between covering common play styles in great detail while leaving out some exotic styles which are rarely observed or are minor variations of the other styles. The isolated roles were: \textit{Carry} - who are usually weak and need protection early on, but are very strong in later stages, often deciding games. Carry typically end up with a high amount of last hits, gold per minute and overall kills, but they can get them in quite different ways.

>>>>

\textit{Dota 2} is team based game in \textit{Multi-player Online Battle Arena} (MOBA). It is divided by two areas, the \textit{Dire} and the \textit{Radiant}. Each team has five players. Every player will choose their hero in pre-match games. Currently, there is 113 heroes, 166 current items. \textit{Dota 2} game design has separated the heroes with three types. The first is agility heroes type, the second is strength heroes type and the last is intelligent heroes type. Every hero has own specialty and weakness. The items can be combined into a upgrade-able item which is will give more status in heroes abilities.

Hero selection is an important aspect of the game. The team needs to balance the heroes with different abilities that able to fulfill certain of roles and strategies. In the beginning of the game, each player will get 650 gold to buy starting items.

\textit{Dota 2} map layout split into three lanes. Each lanes has three defensive towers (see figure \ref{fig:map_lane}). the Dire lanes and the Radiant lanes are quite different. the top of Radiant it is called off-lane, meanwhile, in Dire, the top lane is called safe-lane. The bottom lane of Radiant is called safe-lane, while in Dire it is called off-lane. And the middle lane will remain same.

@@

Player roles or play styles is the way how the player will choose and try to act as their roles or styles in the games. It is important to note that these roles describe a different facet of play than classic player type classifications (e.g. after Bartle \cite{bartle1996hearts}) which aim to classify expressed character traits of the players and were designed to match role-play style games. Comparison of videos, Online guides, and commentary of professional players and commentators are our sources for selection and characterization of the player roles. Player roles definitions and naming conventions will, in any case, differ slightly among the player base and shift over time as the game evolves, constituting another challenge for \textit{Machine Learning} applications in this area. We isolated five player roles for the main \textit{Machine Learning} task, which strike a balance between covering common play styles in great detail while leaving out some exotic styles which are rarely observed or are minor variations of the other styles. The isolated roles were: \textit{Carry} - who are usually weak and need protection early on but are very strong in later stages, often deciding games. Carry typically end up with a high amount of last hits, gold per minute and overall kills, but they can get them in quite different ways. \textit{Mid Laner} - this kind of playing style mostly will get middle lane position since early game. focusing on collecting more gold to obtain high percentage lane of gold in early ten minutes. \textit{Offlaner} - safe lane as their common lane. \textit{Dire} jungle is the safe lane for the Dire team, otherwise, the \textit{Radiant} jungle is the safe lane for the Radiant team. Usually, have a balance between \textit{networth} and efficiently lane gold in early ten minutes. \textit{Roaming Support} - support heroes mostly weak, roaming support will have roaming attribute from replay file, this player role will buy support item such as: \textit{Observer ward} and \textit{Sentry ward} for vision. \textit{Hard Support} - sometimes called the babysitter, because this playing style needs to do everything for supporting the Carry. Mostly, buy support items much more than roaming support.

### 2017-06-06 paper/documents/4.experiment.tex
* fixed grammar

We are using ROpenDota by Rosdyana Kusuma to built our data set. ROpenDota is an API wrapper service for OpenDota in R. We labeling players from replay and the replay based on The International 2016 and Kiev Major, which contains 8419 labeled players. Using replay data from professional players will guarantee our data set from low level data.

@@

Attributes for evaluation is presented in table \ref{table:table1}. While some attributes correspond directly to summary data, attributes that capture positional information the fighting behavior require more complex processing of the replay data. Different attribute filters implemented in the Java library WEKA \cite{hall2009weka} to determine the best set of attributes. Our results with the \textit{WrapperSubsetEval} class with best-first search are presented in table 1. We have chosen this algorithm for our final attribute selection because it resulted in the highest accuracy with our labeled data set. For classification we selected all attributes that were present in at least four folds, excluding assists, as this resulted in the overall highest accuracy. In the following sections we describe the algorithms and heuristics we developed to calculate the attributes that cannot be directly obtained from replay files. They can be grouped into five rough categories: space and movement, early ganks, team fights, support items, and damage types. Not all of the attributes we describe in this section were finally chosen to be used for classification within this work but they might prove valuable for future works.

@@

\textbf{Player Lane} Most of player roles depend on the lane (see figure \ref{fig:map_lane}) which is the most active in the early game. Player lane information also provides the constructor for the other attributes, e.g., the \textit{lane partners}. In the early game, players commonly have three main positions: \textit{top lane}, \textit{mid lane}, or \textit{bottom lane}. In additional, the \textit{jungle} areas also can be used ( see figure \ref{fig:map_jungle}). Or, players can have a \textit{roaming} position, it means that they will move around the map instead of focusing in specific lane.

@@

\textbf{Lane Partners and Solo Lane} To calculate the lane attribute from lane partners and solo lane can be determine by comparing the number of players. Player in \textit{jungle} position or \textit{roaming} are always assigned with zero lane partners, while players who is sharing the other lane, such as top, mid or bottom lane are corresponding with number of their \textit{teammates} in the current lane. Solo lane attribute included in our attributes as an alternative to the lane partners. The value of this attribute will be true if a player is assigned to one of the tree main lanes without any teammate and otherwise false.

@@

\textbf{Damage Types} We categorized damage type based on player roles and requirement of other attributes: damage to tower, damage to heroes, damage to regular creeps, damage to neutral creeps and damage to ancient creeps.

@@

\textbf{Softmax Scaling}, In addition to linear scaling, non linear normalization techniques may be used in cases where data are not evenly distributed around the mean \cite{theodoridis2010introduction}. In such cases the transformations based on nonlinear (i.e., exponential, logarithmic, sigmoid etc) functions can be used to map the data within specified intervals. One such popular technique is so called softmax scaling which squashes the data value non linearly in the interval [0,1]. the normalized feature is given by;

>>>>

We are using ROpenDota by Rosdyana Kusuma to built our data set. ROpenDota is an API wrapper service for OpenDota in R. We labeling players from replay and the replay based on The International 2016 and Kiev Major, which contains 8419 labeled players. Using replay data from professional players will guarantee our data set from low-level data.

@@

Attributes for evaluation is presented in table \ref{table:table1}. While some attributes correspond directly to summary data, attributes that capture positional information the fighting behavior require more complex processing of the replay data. Different attribute filters implemented in the Java library WEKA \cite{hall2009weka} to determine the best set of attributes. Our results with the \textit{WrapperSubsetEval} class with best-first search are presented in table 1. We have chosen this algorithm for our final attribute selection because it resulted in the highest accuracy with our labeled data set. For classification, we selected all attributes that were present in at least four folds, excluding assists, as this resulted in the overall highest accuracy. In the following sections, we describe the algorithms and heuristics we developed to calculate the attributes that cannot be directly obtained from replay files. They can be grouped into five rough categories: space and movement, early ganks, team fights, support items, and damage types. Not all of the attributes we describe in this section were finally chosen to be used for classification within this work but they might prove valuable for future works.

@@

\textbf{Player Lane} Most of the player roles depend on the lane (see figure \ref{fig:map_lane}) which is the most active in the early game. Player lane information also provides the constructor for the other attributes, e.g., the \textit{lane partners}. In the early game, players commonly have three main positions: \textit{top lane}, \textit{mid lane}, or \textit{bottom lane}. In additional, the \textit{jungle} areas also can be used ( see figure \ref{fig:map_jungle}). Or, players can have a \textit{roaming} position, it means that they will move around the map instead of focusing in a specific lane.

@@

\textbf{Lane Partners and Solo Lane} To calculate the lane attribute from lane partners and solo lane can be determined by comparing the number of players. The player in \textit{jungle} position or \textit{roaming} are always assigned with zero lane partners, while players who are sharing the other lane, such as top, mid or bottom lane are corresponding with a number of their \textit{teammates} in the current lane. Solo lane attribute included in our attributes as an alternative to the lane partners. The value of this attribute will be true if a player is assigned to one of the tree main lanes without any teammate and otherwise false.

@@

\textbf{Damage Types} We categorized damage type based on player roles and requirement of other attributes: damage to the tower, damage to heroes, damage to regular creeps, damage to neutral creeps and damage to ancient creeps.

@@

\textbf{Softmax Scaling}, In addition to linear scaling, nonlinear normalization techniques may be used in cases where data are not evenly distributed around the mean \cite{theodoridis2010introduction}. In such cases, the transformations based on nonlinear (i.e., exponential, logarithmic, sigmoid etc) functions can be used to map the data within specified intervals. One such popular technique is so called softmax scaling which squashes the data value non linearly in the interval [0,1]. the normalized feature is given by:

### 2017-06-06 paper/documents/5.conclusion.tex
* fixed grammar

We trained and evaluated several different classifier using 10-fold cross-validation and independent test on our data set. The classifiers was based on existing works and complemented by commonly used classification approaches. It included: \textit{Random forest decision trees}, \textit{logistic regression}, \textit{support vector machines with sequential minimal optimization}, \textit{naive bayes classifier}, and \textit{bayesian networks}. We are using WEKA library and tool for the technical platform. The classifiers were evaluated according to several established performance metric (accuracy, mean absolute error (MAE) \cite{willmott2005advantages}, and area under ROC (AUC) \cite{huang2005using}.

@@

Table \ref{table:table2} list the highest accuracy from 10-fold cross-validation and independent test of our works. Random forest is the most stable with good performance from our works. The accuracy result also reasonable. it means the good accuracy is not over fitting after checking the cross-validation with the independent test results.

@@

Table \ref{table:table3}, comparison result of our works with previous works. Which is our works is quite better in full set classes. Random forest still dominating for the accuracy result.

>>>>

We trained and evaluated several different classifiers using 10-fold cross-validation and independent test on our data set. The classifiers were based on existing works and complemented by commonly used classification approaches. It included: \textit{Random forest decision trees}, \textit{logistic regression}, \textit{support vector machines with sequential minimal optimization}, \textit{naive bayes classifier}, and \textit{bayesian networks}. We are using WEKA library and tool for the technical platform. The classifiers were evaluated according to several established performance metric (accuracy, mean absolute error (MAE) \cite{willmott2005advantages}, and area under ROC (AUC) \cite{huang2005using}.

@@

Table \ref{table:table2} list the highest accuracy from 10-fold cross-validation and independent test of our works. Random forest is the most stable with good performance from our works. The accuracy result also reasonable. it means the good accuracy is not overfitting after checking the cross-validation with the independent test results.

@@

Table \ref{table:table3}, the comparison result of our works with previous works. Which is our works is quite better in full set classes. Random forest still dominating for the accuracy result.

Classify the player roles is not easy. We asking \textit{Dota 2} experts to manually classify the player roles. The responses were highly divergent among them. This illustrates that the classification tasks is difficult, even for human experts. Our data set from the professional tournament (
